---
title: "PRELIMINARY DATA: Real-time SRWC Redd Dewatering Estimates"
author: "BDO Science Division"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
This script constructs real-time winter-run redd dewatering estimates based on data from 

```{r, echo=FALSE}
reddCount <- 1055 
countDate <- "September 12, 2022"
updatedReddInfoDate <- "September 28, 2022"
conservationFish <- 79
#show_extra = TRUE
```

This script constructs real-time winter-run redd dewatering estimates based on most recent data available from CDFW (`r updatedReddInfoDate`). Data are also available in the YYYY Winter-run Data file.xls online at [calfish.org](https://gcc02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fwww.calfish.org%2FProgramsData%2FConservationandManagement%2FCentralValleyMonitoring%2FCDFWUpperSacRiverBasinSalmonidMonitoring.aspx&data=05%7C01%7Clelliott%40usbr.gov%7C689ebb9a6c8243b4f96c08da90f5c542%7C0693b5ba4b184d7b9341f32f400a5494%7C0%7C0%7C637981682646098788%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=A1eQkWPxbkXxnzEvc2K8%2FTmslZ8H8zvxdks3%2F78Yrvw%3D&reserved=0), but these may not reflect the most recent updates.


Please note that all data are preliminary until data collection is finalized. Likewise, there are uncertainties with forecasts which may lead to changes in proposed operations. 


```{r setup, echo=FALSE}
library(tinytex)
library(data.table)
library(plyr)
library(lubridate)

#this function will convert the character values (from xlsx format of numbers with commas) to numeric class values
setClass("num.with.commas")
setAs("character", "num.with.commas", 
        function(from) as.numeric(gsub(",", "", from) ) )

#Read in redd data, converting flow to a numeric and date from character to date value
redds <- read.csv("Redds.csv", colClasses=c('character', 'character', 'num.with.commas'))
redds$Estimated.Date.of.Emergence <- as.Date(redds$Estimated.Date.of.Emergence, "%m/%d/%Y")
reddsAll <- redds

#Read in KES flow data, converting flow to a numeric and date from character to date value
kesFlow <- read.csv("kesFlow.csv")
kesFlow <- read.csv("kesFlow.csv", colClasses=c('character', rep('num.with.commas', length(kesFlow) - 1)))
kesFlow$Date <- as.Date(kesFlow$Date, "%m/%d/%Y")
nAlt <- length(kesFlow) - 1

realFlow <- read.csv("realtimeKES.csv")
realFlow$Date <- as.Date(realFlow$Date, "%m/%d/%Y")

actualDate <- format(Sys.time(), '%d %B, %Y')

```

```{r, echo = FALSE}
# Thousand Acre Feet

for(a in 1:nAlt){
 kesFlow[ ,  paste0("A", substr(names(kesFlow)[a + 1], 4,5), "cfs")] <- NA
 kesFlow[ ,  paste0("A", substr(names(kesFlow)[a + 1], 4,5), "taf")] <- NA
}

for(a in 1:nAlt){
  for(r in 1: nrow(kesFlow)){
      kesFlow[r, paste0("A", substr(names(kesFlow)[a + 1], 4,5), "cfs")] <- kesFlow[r, paste0("Alt", substr(names(kesFlow)[a + 1], 4,5))]
      kesFlow[r, paste0("A", substr(names(kesFlow)[a + 1], 4,5), "taf")] <- (kesFlow[r, paste0("Alt", substr(names(kesFlow)[a + 1], 4,5))]*1.983)/1000
  }
}

septKesFlow <- subset(kesFlow, months(kesFlow$Date) == "September")
octKesFlow <- subset(kesFlow, months(kesFlow$Date) == "October")

```


# Current Redd Count
As of `r countDate`, the unexpanded redd count is **`r reddCount`** Winter-run redds.It is important to note that until data collection is completed for the year this is the **minimum** number of possible redds.  This number will always expand upon final analysis but gives an in-season guard rail of the minimum number of redds this year. From 2018-2021, female expansion has ranged from 0.31 to 1.31 with a 0.7 average, thus we may expect the final number of redds to be closer to `r round(reddCount*1.7, 0)` redds using average expansion on data this year, and using this value, `r 0.01*round(reddCount*1.7, 0)` redds dewatered would be at the 1% take level.

As of `r updatedReddInfoDate`, **`r nrow(subset(reddsAll, reddsAll$Status == "EMERGED")) `** Winter-run redds have **emerged** and **`r nrow(subset(reddsAll, reddsAll$Status == "DEWATERED"))`** have been **dewatered**. This leaves **`r nrow(subset(reddsAll, reddsAll$Status == "OK"))`** shallow water redds of concern.

In addition, **`r conservationFish`** wild adult WRCS were removed to the hatchery as part of conservation actions.

```{r, echo = FALSE}
# Estimated number of redds dewatered
for(a in 1:(nAlt)){
 redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5))] <- NA
}

redds <- subset(redds, redds$Status == "OK" | redds$Status == "DEWATERED")

for(r in 1:nrow(redds)){ #loop over redds
  for(f in 1:nrow(kesFlow)){ #loop over KES flow
    for(a in 1:nAlt){ #loop over alternatives
      if(kesFlow$Date[f] == redds$Estimated.Date.of.Emergence[r]){ #if dates match
        if(kesFlow[f, paste0("Alt", substr(names(kesFlow)[a + 1], 4,5))] <= redds$ACTUAL.or.ESTIMATED..DEWATER.FLOW[r]){ #If KES flow is < or = redd dewatering flow, 1 (dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5))] <- 1
        } else{ #If KES flow is greater than redd dewatering flow, mark 0 (not dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5))] <- redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5))] 
          } #ends else
      }
    } #ends if date is the same
  } #end looping over KES flow dataframe
} #end looping over redds dataframe
#Based on the `r countDate` redd count:
```


```{r, echo = FALSE}
# Estimated number of redds dewatered with a 100cfs buffer
for(a in 1:nAlt){
 redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5), "Buffer")] <- NA
}

for(r in 1:nrow(redds)){ #loop over redds
  for(f in 1:nrow(kesFlow)){ #loop over KES flow
    for(a in 1:nAlt){ #loop over alternatives
      if(kesFlow$Date[f] == redds$Estimated.Date.of.Emergence[r]){ #if dates match
        if(kesFlow[f, paste0("Alt", substr(names(kesFlow)[a + 1], 4,5))] <= redds$ACTUAL.or.ESTIMATED..DEWATER.FLOW[r] + 100){ #If KES flow is < or = redd dewatering flow, 1 (dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5), "Buffer")] <- 1
        } else{ #If KES flow is greater than redd dewatering flow, mark 0 (not dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5), "Buffer")] <- redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5))] 
          } #ends else
      }
    } #ends if date is the same
  } #end looping over KES flow dataframe
} #end looping over redds dataframe
#Based on the `r countDate` redd count:

```


# Table
**Table 1.** Average September Keswick (KES) Flow, monthly flows for September and October (in cfs), total Sept-Oct volume (in TAF), estimated numbers of SRWC redds dewatered, and percent of population that would be lost under each of the proposed alternatives. KES Flow data uses actual flow-to-date as of `r max(realFlow$Date)` and proposed flows for the remainder of the incubation period. Redd dewatering is considered at the actual or estimated dewatering flow and with a 100 cfs buffer applied to the actual/estimated dewatering flow. Percentage of the population lost is based on the `r countDate` count of `r reddCount` Winter-run redds and updated redd counts may be available soon.  
```{r, echo=FALSE, results= TRUE}
Metric <- c("Avg Sept Flow (cfs)", "Avg Oct Flow (cfs)", "Total Sept-Oct Flow (TAF)", "Redds dewatered", "Percent Lost", "Redds dewatered (w/ 100cfs buffer)", "Percent Lost (w/ 100cfs buffer)")
table <- data.frame(Metric)

d <- rep(NA, nAlt)
dflist = vector("list", nAlt) # creating an empty list
dfnames <- rep(NA, nAlt)

for(a in 1:(nAlt)){
 d[a] = paste0("Alt", substr(names(kesFlow)[a + 1], 4,5))
 dfnames[a] = paste0("Alt", substr(names(kesFlow)[a + 1], 4,5))
 names(dflist) = dfnames # naming the list 
 dflist[[a]] <- as.character(c(round(mean(septKesFlow[, paste0("Alt", substr(names(kesFlow)[a + 1], 4,5))], na.rm = TRUE), 0),
                               round(mean(octKesFlow[, paste0("Alt", substr(names(kesFlow)[a + 1], 4,5))], na.rm = TRUE), 0),
                               round(sum(septKesFlow[, paste0("A", substr(names(kesFlow)[a + 1], 4,5), "taf")], na.rm = TRUE) + 
                                       sum(octKesFlow[, paste0("A", substr(names(kesFlow)[a + 1], 4,5), "taf")], na.rm = TRUE), 0),
          sum(redds[ ,paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5))], na.rm = TRUE), 
          round((sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5))], na.rm = TRUE)/reddCount*100), 1), 
          sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5), "Buffer")], na.rm = TRUE), 
          round((sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,5), "Buffer")], na.rm = TRUE)/reddCount*100), 1)))
 table <- cbind(table, dflist[a])
}




knitr::kable(table)
```

# Plots
```{r, echo=FALSE}
reddsDat <- reddsAll
reddsTable <- setDT(reddsDat)[,list(Count=.N), names(reddsDat)]
ymin <- round_any(min(reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW), 1000, f = floor)
ymax <- round_any(max(reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW), 250, f = ceiling)

for(r in 1:nrow(reddsTable)){
  if(reddsTable$Status[r] == "EMERGED"){
    reddsTable$cols[r] <- "light blue"
  } else {
    if(reddsTable$Status[r] == "DEWATERED"){
      reddsTable$cols[r] <- "darkorange2"
  } else {
    reddsTable$cols[r] <- "grey"
    }
  }
}

library(RColorBrewer)
cols <- brewer.pal(nAlt,'Set1')

plot(reddsTable$Estimated.Date.of.Emergence, reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW, 
     pch = 16, col = reddsTable$cols,
     main = "WR Redd Dewatering vs. Flow Alternatives",
     title(xlab = "Emergence Date",  mgp=c(2,1,0)), 
     ylab = "KES Flow (cfs)",
     cex = 2.5, 
     yaxp = c(ymin, ymax, round_any((ymax - ymin)/250, 1, f = ceiling)),
     ylim = c(ymin, ymax),
     xlim = c(min(kesFlow$Date), max(reddsTable$Estimated.Date.of.Emergence) + 5)
     )

grid(nx = NULL, ny = NULL,
     lty = 1,      # Grid line type
     col = "gray", # Grid line color
     lwd = 1)      # Grid line width
abline( v = actualDate, col = "grey", lty = 2, lwd = 1)
#lines(kesFlow$Date, kesFlow$KES_Flow, col = "black", lwd = 2)
text(reddsTable$Estimated.Date.of.Emergence + 0.01, reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW + 10, labels = reddsTable$Count)

for(a in 1:(nAlt)){
  lines(kesFlow$Date , kesFlow[ ,  paste0("A", substr(names(kesFlow)[a + 1], 4,5), "cfs")], col = cols[a], lwd = 2)
}
legend("bottom", inset = c(0, -0.3), legend=c(names(kesFlow)[2:(nAlt+1)], "Emerged", "Dewatered", "Remaining"),
       pch = c( rep(NA, nAlt), 16, 16, 16),
       col=c(cols[1:(nAlt)], "light blue", "darkorange2",  "grey"), 
       lty= c(rep(1, nAlt), NA, NA, NA), cex=0.8, xpd = TRUE, horiz = TRUE)  









#text(x=rep(max(kesFlow$Date)+3, 5), 
 #    y=c(min(kesFlow$Alt1), min(kesFlow$Alt2), 
  #       min(kesFlow$Alt3), min(kesFlow$Alt4), min(kesFlow$Alt5)), 
   #  pos=4, labels=c('red line', 'yellow line', 'orange line', 'green line', 'blue line'))
#data.frame(date = unique(redds$Estimated.Date.of.Emergence), )

```

**Figure 1.** Actual or estimated emergence dates of SRWC redds and actual or estimated dewatering flow for the September-October estimated redd emergence dates as compared to Keswick flow (in cfs) of proposed management alternatives. Points represent emerged, dewatered, or remaining redds. Numbers inside of points indicate how many redds share that estimated emergence date and actual/estimated dewatering flow. Points that fall above/to the right of a flow alternative line are expected to be dewatered given that management alternative is followed. Points that fall below/to the left of/on a flow alternative line are not expected to be dewatered, given that management alternative is followed. Today's date indicated by the dashed line. 

